Some examples in extended logic counting classes

### 2016-06-01 tex/below.tex
6 jun

>>>>

We present a counting class diagram:

\input{classdiag}

Now we will prove the non-trivial inclusions.

For a non-deterministic machine with an output tape $M$, we define $\sp!_M:\{0,1\}^*\to\N$ as follows:
\[
	\sp!_M(x) = \{w \mid \text{every path of $M$ with input $x$ that returns $w$ is an accepting path.} \}
\]
Similarly, we define $\stspanp$ (suggested pronunciation: ``strict span P'') as the class of functions $\sp!_M$ where $M$ is a non-deterministic polynomial time machine.

To prove the inclusion $\spanp \subseteq \shp^{\np}$, we will use a more strong result.

\begin{theo}
	$\spanp \subseteq \stspanp$.
\end{theo}
\begin{proof}
	Let $f \in \spanp$ and let $M$ be the non-deterministic machine for which $f = \sp_M$ that runs in $p(\vert x \vert)$ time. We construct a non-deterministic machine $M'$ that on input $w$ does the following. It simulates $M(x)$, and when the simulation finishes on a final state of $M$, stops. When it finishes on a non-final state, it appends a dummy sequence of characters, for example $0^{p(\vert x \vert)+1}$, to the output. Clearly $f = \sp!_{M'}$, and since $M'$ runs in polynomial time, we have that $f \in \stspanp$.
\end{proof}

\begin{theo}
	$\stspanp = \shp^{\np}$
\end{theo}
\begin{proof}
	To show that $\stspanp \subseteq \shp^{\np}$, consider some $f \in \stspanp$ and some non-deterministic machine $M$ for which $f = \sp!_M$ that runs in $p(\vert x\vert)$. Suppose every output of $M(x)$ has length $p(\vert x\vert)$. Consider the following language in NP:
	\begin{multline*}
		L = \{ (i,M,x,w) \mid M \text{ is a non-determinstic polynomial-time machine and,} \\ \text{either $i = 1$ and $w$ is given as output in some accepting path of $M(x)$}, \\ \text{or $i = 0$ and $w$ is given as output in some rejecting path of $M(x)$}\}.
	\end{multline*}
	 We construct a non-deterministic machine $M'$ with an oracle on $L$ that on input $x$ does the following. Guess some $w\in\{0,1\}^{p(\vert x\vert)}$ Give $(1,M,x,w)$ to the oracle. If it responds {\tt no}, then $M'$ answers {\tt no}. If the oracle responds {\tt yes}, then it proceeds giving $(0,\overline{M},x,w)$ to the oracle, where $\overline{M}$ is the complement of the machine $M$. If the oracle responds {\tt yes}, then $M'$ answers {\tt no}. If it responds {\tt no}, then $M'$ answers {\tt yes}.

	 To show that $\shp^{\np} \subseteq \stspanp$, consider some $f \in \shp^{\np}$, some language $L\in\np$ and some non-deterministic machine $M^L$ for which $f = \acc_{M^L}$. We provide a non-deterministic machine $M'$ as the following non-deterministic procedure. Let $M_L$ the non-deterministic polynomial-time machine machine that computes $L$.
	 \begin{algorithm}
	 \caption{Compute $\acc_{M^L}(x)$ outputs that are accepted in each respective path}
	 \begin{algorithmic}[]
		 \Begin
		 \State $w \gets \varepsilon$.
		 \State Simulate $M^L(x)$:
		 \While{running $M^L(x)$}
		 \State $s\gets$ the current configuration of $M^L(x)$.
		 \State $w \gets w \cdot s$.
		 \If{$s$ is a call to the oracle with input $y$}
		 \State Non-det. simulate $M_L(y)$:
		 \InEach{accepting path} {\bf do} two branches:
		 \State Continue running $M^L(x)$ on $s$ with the state $q_{\texttt{yes}}$.
		 \State Continue running $M^L(x)$ on $s$ with the state $q_{\texttt{no}}$, but rejecting at the end
		 \EndInEach
		 \InEach{rejecting path}
		 \State Continue running $M^L(x)$ on $s$ with the state $q_{\texttt{no}}$.
		 \EndInEach
		 \ElsIf{$\delta(s) \neq \emptyset$} choose from $\delta(s)$
		 \State Continue simulating on that choice.
		 \ElsIf{$s$ corresponds a final state}
		 \State \Return {\tt yes} with output $w$, unless previously decided to be rejected.
		 \Else
		 \State \Return {\tt no} with output $w$.
		 \EndIf
		 \EndWhile
		 \EndBegin
	 \end{algorithmic}
 	 \end{algorithm}

	 Let $x\in\{0,1\}^*$. To show that $\sp!_{M'}(x) = \acc_{M^L}(x)$, first note that for each strictly accepting output $w$ of $M'(x)$ there exists some accepting path $p$ of $M^L(x)$ such that $p = w$. Now note that for each accepting path $p$ of $M^L(x)$ there exists an output $w$ of $M'(x)$ such that $p = w$. This output is clearly accepted in some branch. To show that this output is not rejected in any branch, by contradiction suppose that it is. This branch cannot correspond to a rejecting path of $M^L(x)$ and then it must be rejected on line 11, from simulating a negative answer of the oracle in one of its accepting paths. This branch corresponds to a path that is never reachable since in $M^L(x)$ the oracle answered {\tt yes}, and we follow to a contradiction. 
\end{proof}
